Tight Bounds for Dynamic Balanced Graph Partitioning

### 2020-05-09 paper/disc20clustering.tex
isolated double definition fix
PPL proof

\noindent
\textbf{Perfect Partition Learner algorithm.}
Fix the initial partition
$P_I = I_1, \dots, I_{\ell}$ and \OPT's final partition
$P_F = F_1, \dots, F_{\ell}$.
The \emph{distance} of a partition $P = C_1, \dots, C_{\ell}$ from the initial partition,
defined as 
$\Delta(P, P_I) := \sum_{j=1}^{\ell} | C_j \setminus I_j |$,
 is the number of nodes in $P$ that do not reside in their initial cluster.
In other words,
at least $\Delta(P, P_I)$ node migrations are required in order to reach the partition $P$ from $P_I$, and thus
$\OPT \geq \Delta^*:= \Delta(P_F, P_I) $.
We  use $ \Delta(P_F) $ wherever the initial partition is clear from the context.
The scheme of the algorithm can be found in the Algorithm \ref{alg:PPL}.

@@

\begin{property} \label{prop:dist<OPT}
	Let $P$ be any configuration chosen by \PPL at any time.
	Then, $\Delta(P,P_I) \leq \Delta^*$.	
\end{property}

\begin{lemma}	\label{lemma:rebalancecost}
	The cost of each repartitioning by \PPL is $2\cdot\OPT$.
\end{lemma}
\begin{proof}
	Consider the repartitioning that transforms $P_{t-1}$ to $P_t$ upon the request $\sigma_t$.
	Let $M \subset V$ denote the set of nodes that migrate during this process.
	Let $M^-$ and $M^+$ denote the subset of nodes that (respectively)
	enter or leave their original cluster during the repartitioning.    
	Then,
	$M = M^+ \cup M^-$.
	Since \maciek{at least?} $|M^-|$ nodes are not in their original cluster before the repartitioning (i.e., in $P_{t-1}$),
	the distance before the repartitioning is $\Delta(P_{t-1},P_I) \geq | M^-|$.
	Analogously,
	the distance afterwards is $\Delta(P_{t},P_I) \geq | M^+|$.
	Thus,
	$|M| \leq \Delta(P_{t-1},P_I) + \Delta(P_{t},P_I)$.
	By Property \ref{prop:dist<OPT},
	$\Delta(P_{t-1},P_I) , \Delta(P_{t},P_I) \leq \Delta^* \leq \OPT$
	and thereby we have	
	$|M| \leq 2\cdot\OPT$.
	\maciek{the comma between $\Delta$ is confusing. BOTH are bounded? If so, expand the sentence and write both individual bounds.}
\end{proof}

\begin{theorem}	\label{thm:upperbound}
	\PPL reaches the final configuration $P_F$ and it is $(2\cdot k\cdot\ell)$-competitive.
	\maciek{Do not hide that we have $2(k-1)\ell$ in the theorem. This is a better bound.}
\end{theorem}
\begin{proof}
	On each inter-cluster request,
	the algorithm enumerates all $\ell$-way partitions of components
	that are in the same (closest) distance of $P_I$.
	That is, 
	once it reaches a configuration $P$ at distance $\Delta^* = \Delta(P, P_I)$,
	it does not move to a configuration
	$P', \Delta(P', P_I) > \Delta^*$,
	before it enumerates all configurations at distance $\Delta^*$.
	Therefore,
	\PPL eventually reaches $\Delta^*=\OPT$ and the configuration $P_F$.
	%    including the request that completes revealing of all components that are collocated in $P_F$.
	There are at most $(k-1)\cdot\ell < k\cdot\ell $ calls   to $\mathit{repartition}$
	(i.e., the number of internal edges in $P_F$).
	\maciek{Clarify why the number of internal edges bound the number of repartitions.}
	By Lemma \ref{lemma:rebalancecost},
	each repartition costs at most $2\cdot\OPT$.
	The total cost is therefore at most $2\cdot\OPT\cdot k\cdot\ell$, which implies the competitive ratio.
	%    \mahmoud{This is the cost of moving and the cost of remote comm. is not counted.
	%    	So it is 4-competitive (?)}
\end{proof}

>>>>

\noindent
\textbf{Perfect Partition Learner algorithm.}
Fix the initial partition
$P_I :=\{ I_1, \dots, I_{\ell}\}$ and \OPT's final partition
$P_F := \{F_1, \dots, F_{\ell}\}$.
The \emph{distance} of a partition
 $P = \{C_1, \dots, C_{\ell}\}$ from the initial partition,
defined as 
$\Delta(P, P_I) := \sum_{j=1}^{\ell} | C_j \setminus I_j |$,
 is the number of nodes in $P$ that do not reside in their initial cluster.
In other words,
at least $\Delta(P, P_I)$ node migrations are required in order to reach the partition $P$ from $P_I$, and thus
$\OPT \geq \Delta^*:= \Delta(P_F, P_I) $.
We  use $ \Delta(P_F) $ wherever the initial partition is clear from the context.
The scheme of the algorithm can be found in the Algorithm \ref{alg:PPL}.

@@

\begin{property} \label{prop:dist<OPT}
	Let $P$ be any partition chosen by \PPL at any time.
	Then, $\Delta(P) \leq \Delta^*$.	
\end{property}

\begin{lemma}	\label{lemma:rebalancecost}
	The cost of each repartitioning by \PPL is $2\cdot\OPT$.
\end{lemma}
\begin{proof}
	Let $P_i$ denote the partition of \PPL immediately after serving $\sigma_{i}$.
	Consider the repartitioning that transforms 
	$P_{t-1}$ to $P_t$ upon the request $\sigma_t$.
	Let $M \subset V$ denote the set of nodes that migrate during this process.
	Let $M^-$ and $M^+$ denote the subset of nodes that, respectively,
	enter or leave their initial cluster during the repartitioning.    
	Then,
	$M = M^+ \cup M^-$.
	Since at least $|M^-|$ nodes are not in their initial cluster before the repartitioning (i.e., in $P_{t-1}$),
	the distance before the repartitioning is $\Delta(P_{t-1}) \geq | M^-|$.
	Analogously,
	the distance afterwards is $\Delta(P_{t}) \geq | M^+|$.
	Thus,
	$|M| \leq \Delta(P_{t-1}) + \Delta(P_{t})$.
	By Property \ref{prop:dist<OPT},
	$\Delta(P_{t-1})  \leq \Delta^*$ and
	$\Delta(P_{t}) \leq \Delta^*$.
	Since $\Delta^* \leq \OPT$,
	we obtain	 $|M| \leq 2\cdot\OPT$.
\end{proof}

\begin{theorem}	\label{thm:upperbound}
	\PPL reaches the final partition $P_F$
	 and it is $(2\cdot (k-1)\cdot\ell)$-competitive.
\end{theorem}
\begin{proof}
	On each inter-cluster request,
	the algorithm enumerates all $\ell$-way partitions of components
	that are in the same (closest) distance to $P_I$.
	That is, 
	once it reaches a partition $P$ at distance $\Delta^* = \Delta(P)$,
	it does not move to a partition
	$P', \Delta(P') > \Delta^*$,
	before it enumerates all partitions at distance $\Delta^*$.
	Therefore,
	\PPL eventually reaches  the partition $P_F$ at distance $\Delta^*=\OPT$.
	With each distinct request, 
	the size of some component increases by one.
	For any cluster $F_i \in P_F$,
	we have $\sum_{C \in F_i} |C| = k$.
	A component $C \in F_i$,
	 initially begins as an isolated node and it grows by gaining $|C|-1$ more nodes.
	Hence, the total number of times a component in $F_i$ grows is 
	$\sum_{C \in F_i} (|C|-1) \leq k-1$.
	Therefore, there are at most  $(k-1)\cdot\ell $ distinct requests
	for which \PPL performs a repartitioning.
	I.e.,
	 \PPL performs at most $(k-1)\cdot\ell $ repartitionings.
	By Lemma \ref{lemma:rebalancecost},
	each repartitioning costs at most $2\cdot\OPT$.
	The total cost is therefore at most $2\cdot\OPT\cdot (k-1) \cdot\ell$, which implies the competitive ratio.
\end{proof}
